Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-15 talk/pepm2011/paper.tex
added reference to constructor specialisation
some minor changes + requeusts

@@

While we do not offer a formal proof of it, it should be relatively clear
that the algorithm presented above is sound: it works by delaying (and
often completely removing) some operations.  The algorithm runs in a
single pass over the list of operations.  We can check that altough
recursively lifting a static object is not a constant-time operation,
the algorithm only takes a total time linear in the length of the trace.
Moreover, it gives the ``best'' possible result within its constrains,
\eg in term of the number of residual operations.  The
algorithm itself is not particularly complex or innovative; our focus is
rather that \emph{in the context of tracing JITs} it is possible to find a
simple enough algorithm that still gives the best results.

@@

One problem to the successful application of the allocation removal
techniques described in the previous sections is the presence of
frame-introspection features in many dynamic languages and the fact
that frames are heap allocated. Languages such as Python and Smalltalk
allow the programmer to get access to the frames object that the
interpreter uses to store local variables. This is a useful feature,
as it makes the implementation of a debugger possible in Python
without needing much explicit support from the VM level. On the other
hand, it severely hinders the effectiveness of allocation removal,
because every time an object is stored into a local variable, it is
stored into the frame-object, which makes it escape.

@@

Partially known data structures are built directly into Prolog (via unbound
logic variables) and thus the treatment of partially static data structures
was part of partial evaluation of Prolog programs from the early stages
\cite{lloyd_partial_1991}. One effect of unfolding in Prolog is that terms that
are constructed and immediately matched again, completely disappear in the
residual program. This is similar to what our optimization does for an
imperative language.

xxx:
partially static data structures: kenichi asai's thesis?

xxx: 
Other related work is compile-time garbage collection \cite{mazur_practical_2001}, which
tries to ....;
  separation logic; John Hughes: type specialization

>>>>

XXX Apply the rules of the algorithm to parts of the sample trace

@@

While we do not offer a formal proof of it, it should be relatively clear
that the algorithm presented above is sound: it works by delaying (and
often completely removing) some operations.  The algorithm runs in a
single pass over the list of operations.  We can check that altough
recursively lifting a static object is not a constant-time operation,
the algorithm only takes a total time linear in the length of the trace.
Moreover, it gives the ``best'' possible result within its constraints,
\marginpar{Why? What are the constraints ? Justify}
\eg in term of the number of residual operations.  The
algorithm itself is not particularly complex; our focus is
rather that \emph{in the context of tracing JITs} it is possible to find a
simple enough algorithm that still gives the best results.

@@

XXX unclear what the relationship is to the technique; has it been used in the experiments??
XXX Maybe move to discussion and related work section?

One problem to the successful application of the allocation removal
techniques described in the previous sections is the presence of
frame-introspection features in many dynamic languages and the fact
that frames are heap allocated. Languages such as Python and Smalltalk
allow the programmer to get access to the frames object that the
interpreter uses to store local variables. This is a useful feature,
as it makes the implementation of a debugger possible in Python
without extra support from the VM. On the other
hand, it severely hinders the effectiveness of allocation removal,
because every time an object is stored into a local variable, it is
stored into the frame-object, which makes it escape.

@@

Partially known data structures are built directly into Prolog (via unbound
logic variables) and thus the treatment of partially static data structures
was part of partial evaluation of Prolog programs from the early stages
\cite{lloyd_partial_1991}. One effect of unfolding in Prolog is that terms that
are constructed and immediately matched again, completely disappear in the
residual program. This is similar to what our optimization does for an
imperative language.
In functional programming this idea was introduced as
 constructor specialisation by Mogensen \cite{Mogensen:PEPM93}.)

%xxx:
%partially static data structures: kenichi asai's thesis?

xxx: 
Other related work is compile-time garbage collection \cite{mazur_practical_2001}, which
tries to ....;
  separation logic
  %; John Hughes: type specialization
